---
title: "Preprocessing network Zurich"
author: "Nicola Maiani"
date: "5/2/2023"
output:
  html_document:
    df_print: paged
    pandoc_args: [
      "+RTS", "-K100000000000000",
      "-RTS"
    ]
---

# Processing network
## Preliminaries
```{r, message = FALSE, warning = FALSE}
test_pkg <- function(x)
  {
    if (!require(x, character.only = TRUE))
    {
      install.packages(x, dep = TRUE)
        if(!require(x, character.only = TRUE)) stop("Package not found")
    }
}

test_pkg("sf")
test_pkg("ggplot2")
test_pkg("sfnetworks")
test_pkg("tidyverse")
test_pkg("tidygraph")
test_pkg("dplyr")
test_pkg("tmap")
test_pkg("igraph")
test_pkg("units")
test_pkg("stplanr")

dataFolder <- here::here("data")
RFolder <- here::here()
figureFolder <- here::here("figs")
```

## Load data
```{r}
# City borders
communities <- st_read(file.path(dataFolder, "gemeindegrenzen/swissBOUNDARIES3D_1_4_TLM_HOHEITSGEBIET.shp")) # load Swiss communities
border <- communities %>% filter(OBJEKTART == "Gemeindegebiet" & NAME == "Zürich") %>% st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 2056)

# Population data
statpop_ch <- read.csv(file.path(dataFolder, "STATPOP2021/STATPOP2021.csv"), sep = ";")
statpop_ch <- st_as_sf(x = statpop_ch, coords = c("E_KOORD", "N_KOORD"), crs = 2056)
statpop_ch <- statpop_ch %>% select(B21BTOT)
statpop <- st_filter(statpop_ch, border)

# Network
network <- st_read(file.path(dataFolder, "network_ch/network_ch.shp")) # load Swiss OSM network
network <- st_transform(network, 2056)
network <- st_filter(network, border, .pred = st_intersects) # filter network to city borders

# Badis in Zurich as test set
badis_zh <- st_read(file.path(dataFolder, "zurich/badis.shp"))
badis_zh <- st_transform(badis_zh, 2056)
```

## Frequency of attribute values of street segments
```{r}
network %>% group_by(highway) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(route) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(junction) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(motorroad) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(access_con) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(barrier) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(bicycle) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(bicycle_co) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(bicycle_ro) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(bridge) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(cyclestree) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(cycleway_b) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(cycleway_l) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(cycleway_r) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(cycleway_1) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(cycleway_2) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(oneway_bic) %>% count() %>% arrange(desc(n))
```

```{r}
network %>% group_by(ramp_bicyc) %>% count() %>% arrange(desc(n))
```

```{r}
# Define function to summarize network topology
summary_net <- function(net){
  # net <- net %E>%
  #   mutate(
  #     weight = unclass(edge_length()))
  # 
  # net <- net %E>%
  # mutate(
  #   weight = ifelse(weight == 0, 0.01, weight)
  # )
  # 
  # net <- net %N>% mutate(
  #   degree = centrality_degree(weights = NULL),
  #   closeness = centrality_closeness(weights = weight),
  #   betweenness = centrality_betweenness(directed = FALSE, weights = weight)
  # )
  # 
  # dual_morph <- net %>%
  #   morph(to_linegraph)
  # dual_net <- dual_morph$line_graph
  # 
  # dual_net <- dual_net %N>% mutate(
  #   degree = centrality_degree(weights = NULL),
  #   closeness = centrality_closeness(weights = NULL),
  #   betweenness = centrality_betweenness(weights = NULL)
  # )
  
  cat(paste("Number of nodes:", nrow(net %>% activate(nodes) %>% st_as_sf())), sep = "\n")
  cat(paste("Number of edges:", nrow(net %>% activate(edges) %>% st_as_sf())), sep = "\n")
  cat(paste("Number of components:", count_components(net, mode = "weak")), sep = "\n")
  cat(paste("Number of under- or overpassing edges:", nrow(net %>% activate(edges) %>% filter(edge_crosses(.E())) %>% st_as_sf())), sep = "\n")
  cat(paste("Number of multiple edges:", sum(which_multiple(net))), sep = "\n")
  cat(paste("Number of loop edges:", sum(which_loop(net))), sep = "\n")
  
  # cat(paste("Node centralities:"), sep = "\n")
  # print(net %N>% data.frame() %>% select(degree, closeness, betweenness) %>% summary())
  # 
  # cat(paste("Edge centralities:"), sep = "\n")
  # print(dual_net %N>% data.frame() %>% select(degree, closeness, betweenness) %>% summary())
}
```

## Create sfnetwork
```{r}
# Create undirected sfnetwork
network <- as_sfnetwork(network, directed = FALSE)

# Make edges spatially explicit
network <- tidygraph::convert(
  network,
  to_spatial_explicit,
  .clean = TRUE
)
```

```{r}
# Subdivide edges for topological correctness
full_network <- convert(network, to_spatial_subdivision)
# summary_net(full_network)
```

### Centralities of full/unfiltered network
```{r}
# Use edge length as weight and calculate centralities
full_network <- full_network %E>%
  mutate(
    weight = edge_length()
  )

full_network <- full_network %N>% mutate(
    degree = centrality_degree(weights = NULL),
    closeness = centrality_closeness(weights = weight),
    betweenness = centrality_betweenness(weights = weight, directed = FALSE)
  )

dual_morph <- full_network %>%
  morph(to_linegraph)
dual_net <- dual_morph$line_graph

dual_net <- dual_net %N>% mutate(
  degree = centrality_degree(weights = NULL),
  closeness = centrality_closeness(weights = NULL),
  betweenness = centrality_betweenness(weights = NULL)
  )

dual_net_df <- dual_net %N>% data.frame()

full_network <- full_network %E>%
  mutate(
    degree_edge = dual_net_df$degree,
    closeness_edge = dual_net_df$closeness,
    betweenness_edge = dual_net_df$betweenness
  )
```

```{r}
# Save full network
if (!file.exists(here::here("data", "full_network.rds"))) {
saveRDS(full_network, file = here::here("data", "full_network.rds"))
}
```

```{r}
# Load full network with centralities
full_network <- readRDS(here::here("data", "full_network_zh.rds")) # needs to be converted to sfnetwork!
```

### Filter with respect to attributes
The filtering has been done with taginfo Switzerland and visually assessing values (if it makes sense or not). Use full network as first input!
```{r}
full_network %E>% as_tibble(spatial = FALSE) %>% group_by(highway) %>% count() %>% arrange(desc(n))
relevant_highway <- c("cycleway","living_street","path","primary","primary_link","pedestrian","residential","road","secondary","secondary_link","service","tertiary","tertiary_link","track","unclassified","yes","mini_roundabout", "crossing")
```

```{r}
# Filter with respect to access and barriers
network <- full_network %E>% # use full network here!
  filter(is.na(access) | access %in% c("yes","customers","destination","permissive","unknown","public","residents","24/7","residential") | bicycle %in% c("yes","designated","permissive","private","destination","unknown")) %N>% filter(!node_is_isolated())
  
network <- network %E>%
  filter(is.na(barrier)) %N>% filter(!node_is_isolated())

# summary_net(network)
```

```{r}
# Filter with respect to highways, also include other types of highways if bicycle explicitly allowed
network <- network %E>%
  filter(highway %in% relevant_highway | bicycle %in% c("yes","designated","permissive","private","destination","unknown") | footway %in% c("crossing")) %N>% filter(!node_is_isolated())

network <- network %E>%
  filter(!(highway %in% c("construction"))) %N>% filter(!node_is_isolated())

network <- network %E>%
  filter(is.na(motorroad)) %N>% filter(!node_is_isolated())

# summary_net(network)
```

```{r}
network <- network %E>%
  filter(is.na(ramp_bicyc) | ramp_bicyc == "yes") %N>% filter(!node_is_isolated())

# summary_net(network)
```

```{r}
network <- network %E>%
  filter(is.na(route) | route == "bicycle") %N>% filter(!node_is_isolated())

# summary_net(network)
```

```{r}
network <- network %E>%
  filter(!(vehicle %in% c("agricultural","forestry","no","permit","delivery","agricultural;forestry")) | is.na(vehicle) | bicycle %in% c("yes","designated","permissive","private","destination","unknown")) %N>% filter(!node_is_isolated())

# summary_net(network)
```

```{r}
network <- network %E>%
  filter(!(bicycle %in% c("dismount","mtb","no","use_sidepath"))) %N>% filter(!node_is_isolated())
# summary_net(network)
```

```{r}
network <- network %E>%
  filter(!(smoothness %in% c("very_bad","impassable","horrible","very_horrible"))) %N>% filter(!node_is_isolated())
```

### Join important footways to network
It has been tried to join footways important for the connectivity back to the network. However, approach was dismissed, since results were not satisfying and
footways in OSM should be exclusively for pedestrians. 
```{r}
# # Select important footways
# footways <- full_network %E>% filter(highway == "footway") %E>% filter(footway != "sidewalk" | is.na(footway)) %E>% 
#   filter(is.na(access) | access %in% c("yes","customers","destination","permissive","unknown")) %E>%
#   filter(!(vehicle %in% c("agricultural","forestry","no","permit"))) %E>%
#   filter(!(bicycle %in% c("yes","designated","permissive","private","destination")) | is.na(bicycle)) %N>% 
#   filter(!node_is_isolated()) # get footways, which aren't sidewalks
# 
# footways_df <- footways %E>% st_as_sf() %>% data.frame()
# summary(footways_df$betweenness_edge)
# 
# footways_important <- footways %E>% filter(betweenness_edge > median(footways_df$betweenness_edge)) %N>% 
# #  filter(drop_units(weight) < 150) %N>% 
#   filter(!node_is_isolated())
# 
# tmap_mode("view") # set to interactive mode
# tm_tiles("CartoDB.Positron") +
# tm_shape(footways_important %E>% st_as_sf(), "footways") +
#   tm_lines() + tm_shape(st_as_sf(footways_important %>% activate("nodes"), "nodes")) + tm_dots()
```

```{r}
# # Join them to network
# network <- st_network_join(network, footways_important)
```

```{r}
# # Use edge length as weight and calculate centralities on filtered network
network <- network %E>%
  mutate(
    weight = edge_length()
  )
# 
# network <- network %N>% mutate(
#     degree = centrality_degree(weights = NULL),
#     closeness = centrality_closeness(weights = weight),
#     betweenness = centrality_betweenness(weights = weight, directed = FALSE)
#   )
# 
# dual_morph <- network %>% 
#   morph(to_linegraph)
# dual_net <- dual_morph$line_graph
#   
# dual_net <- dual_net %N>% mutate(
#   degree = centrality_degree(weights = NULL),
#   closeness = centrality_closeness(weights = NULL),
#   betweenness = centrality_betweenness(weights = NULL)
#   )
```

```{r}
# # Plot degree centrality of nodes: Degree centrality not correct, since topology not correct => need to subdivide edges
# tmap_mode("view") # set to interactive mode
# tm_tiles("CartoDB.Positron") +
# tm_shape(st_as_sf(network %>% activate("edges")), "edges") +
#   tm_lines() +
# tm_shape(st_as_sf(network %>% activate("nodes"), "nodes")) +
#   tm_dots(col = "degree")
```

```{r}
# Short summary of network topology
# summary_net(network)
```

## Preprocessing topology
```{r}
# Subdivide edges: Already done above
network <- convert(network, to_spatial_subdivision)
```

```{r}
# Plot network with subdivided edges with degree centrality
network <- network %N>% mutate(
    degree = centrality_degree(weights = NULL))

tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(network %>% activate("edges")), "edges") +
  tm_lines() +
tm_shape(st_as_sf(network %>% activate("nodes"), "nodes")) +
  tm_dots(col = "degree")
```

```{r}
net_df <- network %E>% data.frame()

net_df %>% group_by(highway) %>% count() %>% arrange(desc(n))
net_df %>% group_by(route) %>% count() %>% arrange(desc(n))
net_df %>% group_by(junction) %>% count() %>% arrange(desc(n))
net_df %>% group_by(bridge) %>% count() %>% arrange(desc(n))
```

```{r}
# Remove pseudo nodes, only if the following attributes are the same
equal_pseudo <- c("access","bicycle","bridge","cyclestree","cycleway","cycleway_b","cycleway_l","cycleway_1","cycleway_r","cycleway_2","foot","footway","segregated","highway","maxspeed","oneway","oneway_bic","ramp_bicyc","traffic_ca","tunnel","vehicle")

# First the NA values have to be replaced, in order to evaluate equality in the smoothing function
network <- network %E>%
  dplyr::mutate(maxspeed = as.numeric(maxspeed),
                cyclestree = as.character(cyclestree))
```

```{r}
# First the NA values have to be replaced, in order to evaluate equality in the smoothing function
network <- network %E>% mutate_if(is.character, ~replace_na(., "unknown"))
network <- network %E>% mutate_if(is.numeric, ~replace_na(., -99))

# Define how the attributes of merged segments should be combined
combinations = list(
  osm_id = "first",
  access = "first",
  barrier = "first",
  bicycle = "first",
  bridge = "first",
  cyclestree = "first",
  cycleway = "first",
  cycleway_b = "first",
  cycleway_l = "first",
  cycleway_1 = "first",
  cycleway_r = "first",
  cycleway_2 = "first",
  foot = "first",
  footway = "first",
  segregated = "first",
  highway = "first",
  junction = "first",
  maxspeed = "first",
  motorroad = "first",
  oneway = "first",
  oneway_bic = "first",
  ramp_bicyc = "first",
  traffic_ca = "first",
  tunnel = "first",
  vehicle = "first",
  width = "first",
  weight = "sum",
  "ignore" # rest of attributes is ignored and dropped
)

network <- network %>% convert(to_spatial_smooth, require_equal = equal_pseudo, summarise_attributes = combinations, .clean = TRUE)
```

```{r}
# Look at multiple edges
nrow(network %E>% filter(edge_is_multiple()) %>% st_as_sf())

network %E>% filter(edge_is_multiple()) %>%
  mutate(
    length = edge_length()
  ) %>% as_tibble() %>% select(length) %>% summary()

tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(network %E>% filter(edge_is_multiple())), "edges") + tm_lines()
```

```{r}
# Remove multiple edges: Removes multiple edges, but also loop edges
network <- network %E>%
  arrange(edge_length()) # order in ascending order

# Define how attributes should be combined
combinations = list(
  osm_id = "first",
  access = "first",
  access_con = "first",
  barrier = "first", 
  bicycle = "first",
  bicycle_co = "first",
  bicycle_ro = "first",
  bridge = "first", 
  cyclestree = function(x) if ("yes" %in% x) "yes" else NA,
  cycleway = function(x){
    if("track" %in% x | "yes" %in% x){
      "track"
    } else if ("opposite_lane" %in% x | "lane" %in% x){
      "lane"
    } else if ("share_busway" %in% x){
      "share_busway"
    } else {
      NA
    }
  },
  cycleway_b = function(x){
    if("track" %in% x | "yes" %in% x){
      "track"
    } else if ("opposite_lane" %in% x | "lane" %in% x){
      "lane"
    } else if ("share_busway" %in% x){
      "share_busway"
    } else {
      NA
    }
  },
  cycleway_l = function(x){
    if("track" %in% x | "yes" %in% x){
      "track"
    } else if ("opposite_lane" %in% x | "lane" %in% x){
      "lane"
    } else if ("share_busway" %in% x){
      "share_busway"
    } else {
      NA
    }
  },
  cycleway_r = function(x){
    if("track" %in% x | "yes" %in% x){
      "track"
    } else if ("opposite_lane" %in% x| "lane" %in% x){
      "lane"
    } else if ("share_busway" %in% x){
      "share_busway"
    } else {
      NA
    }
  },
  cycleway_1 = "first",
  cycleway_2 = "first",
  foot = "first",
  footway = "first",
  segregated = "first",
  highway = function(x) if ("cycleway" %in% x) "cycleway" else x[1],
  junction = "first",
  maxspeed = "min",
  motorroad = "first", 
  mob = "first",
  oneway = "first",
  oneway_bic = "first",
  ramp_bicyc = "first",
  route = "first",
  smoothness = "first", 
  traffic_ca = "first",
  tunnel = "first", 
  vehicle = "first",
  width = "first", 
  weight = "first",
  z_order = "first", 
  "ignore"
)

network <- network %E>% convert(to_spatial_simple, summarise_attributes = combinations) # choose shortest connection
```

```{r}
# Look at loop edges
nrow(network %E>% filter(edge_is_loop()) %>% st_as_sf())
loops <- network %E>% filter(edge_is_loop())
loop_edges <- loops %E>% as_tibble()
loop_edges$weight


tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(network %E>% filter(edge_is_loop())), "edges") + tm_lines()
```

```{r}
# Drop columns and calculate degree centrality of nodes on new network
network <- network %E>% select(-any_of(c("osm_id", "access_con", "bicycle_co", "bicycle_ro", "mob", "route", "smoothness", "z_order", "way_area", ".tidygraph_edge_index", "degree_edge", "closeness_edge", "betweenness_edge"))) %N>%
  select(-any_of(c("degree", "closeness", "betweenness")))

network <- network %N>% mutate(
    degree = centrality_degree(weights = NULL))

summary_net(network)
```

Most of the crossings included are redundant, since they connect sidewalks, which are not included in the bike network. Let's therefore only include important crossings using the degree centrality of the starting and end nodes. 
```{r}
# Get rid of redundant crossings in the network
nodes_tibble <- network %N>% as_tibble(spatial = FALSE) %>% select(-geometry)
edges_tibble <- network %E>% as_tibble(spatial = FALSE)
edges_tibble <- left_join(edges_tibble, nodes_tibble, by = c("from" = ".tidygraph_node_index")) %>% rename(from_degree = degree)
edges_tibble <- left_join(edges_tibble, nodes_tibble, by = c("to" = ".tidygraph_node_index")) %>% rename(to_degree = degree)

network <- network %E>%
  mutate(from_degree = edges_tibble$from_degree,
         to_degree = edges_tibble$to_degree,
         keep_edge = ifelse((footway == "crossing" | highway == "crossing") & (from_degree == 1 | to_degree == 1), 0, 1))

network <- network %E>%
  filter(keep_edge == 1) %N>% 
  filter(!node_is_isolated())
```

```{r}
# Plot degree centrality of nodes
tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(network %>% activate("edges")), "edges") +
  tm_lines() +
tm_shape(st_as_sf(network %>% activate("nodes"), "nodes")) +
  tm_dots(col = "degree")
```

```{r}
# Main component
components <- components(network)$membership
main_comp_id <- which.max(components(network)$csize)

network <- network %N>%
  mutate(
    component = components
  )

main_graph <- network %N>%
  filter(component == main_comp_id)

summary_net(main_graph)
```

```{r}
# Plot main connected component
tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(main_graph %>% activate("edges")), "edges") +
  tm_lines() +
tm_shape(st_as_sf(main_graph %>% activate("nodes"), "nodes")) +
  tm_dots()
```

```{r}
# Plot main connected component with ggplot()
ggplot() + geom_sf(data = main_graph %E>% st_as_sf(), fill="white", size=0.45) +
  coord_sf(datum = 2056) +
  theme_void() +
  labs(title="Bike network")
```

```{r}
# Sub-components
sub_graphs <- network %N>%
  filter(component != main_comp_id)

sub_graphs <- sub_graphs %N>% 
  mutate(
    nindex = 1:nrow(sub_graphs %N>% as_tibble())
  )

sub_graphs_edges <- sub_graphs %E>% as_tibble() %>% as.data.frame()
sub_graphs_nodes <- sub_graphs %N>% as_tibble() %>% as.data.frame() %>% select(nindex, component)
sub_graphs_edges <- sub_graphs_edges %>% left_join(sub_graphs_nodes, by = c("from"="nindex"))
sub_graphs_edges$weight <- as.numeric(sub_graphs_edges$weight)
sub_graphs_edges %>% group_by(component) %>% summarise(tot_length = sum(weight)) %>% summary()

summary_net(sub_graphs)
```

If the disconnected sub-components are small in length and more or less randomly distributed in the study area, just work with extracted main connected component. 
```{r}
# Plot disconnected sub-components
tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(sub_graphs %>% activate("edges")), "edges") +
  tm_lines() +
tm_shape(st_as_sf(sub_graphs %>% activate("nodes"), "nodes")) +
  tm_dots()
```

```{r}
# Check for under- and overpassing edges
nrow(main_graph %>% activate(edges) %>% filter(edge_crosses(.E())) %>% st_as_sf())
underoveredges <- main_graph %>% activate(edges) %>% filter(edge_crosses(.E())) %N>% filter(!node_is_isolated())

underoveredges %E>% as_tibble() %>% group_by(bridge) %>% count() %>% arrange(desc(n))
underoveredges %E>% as_tibble() %>% group_by(tunnel) %>% count() %>% arrange(desc(n))

bridges <- underoveredges %E>% filter(bridge == "yes" | bridge == "viaduct") %N>% filter(!node_is_isolated())

tunnels <- underoveredges %E>% filter(tunnel == "yes") %N>% filter(!node_is_isolated())
no_bridge_tunnel <- underoveredges %E>% filter(bridge == "unknown" & tunnel == "unknown") %N>% filter(!node_is_isolated())

tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(bridges %E>% st_as_sf(), "bridges") +
  tm_lines(col = "red") + tm_shape(tunnels %E>% st_as_sf(), "tunnels") + tm_lines(col = "blue") +
  tm_shape(no_bridge_tunnel %E>% st_as_sf(), "no") + tm_lines(col = "black")
```

```{r}
# Use main graph as final network
network <- main_graph

# Assign edge length to edges of network
network <- network %E>%
  mutate(
    weight = edge_length())

network %E>% as_tibble(spatial = FALSE) %>% dplyr::select(weight) %>% summary()
net_df <- network %E>% as_tibble(spatial = FALSE)
quantile(net_df$weight, 0.99)
```

```{r}
# Manipulate maxspeed columns to numeric
network <- network %E>%
  dplyr::mutate(cyclestree = as.character(cyclestree))

network <- network %E>% mutate_if(is.character, ~replace_na(., "unknown"))
network <- network %E>% mutate_if(is.numeric, ~replace_na(., -99))

# Drop columns used to look at crossings
network <- network %E>% select(-any_of(c("from_degree", "to_degree", "keep_edge")))
```

```{r}
# Define function to add edges between two nodes
addedge <- function(node1_index, node2_index, highway){
  n1 <- network %N>% filter(.tidygraph_node_index == node1_index) %>% st_coordinates()
  n2 <- network %N>% filter(.tidygraph_node_index == node2_index) %>% st_coordinates()
  l <- st_linestring(rbind(c(n1[, 1], n1[, 2]), c(n2[, 1], n2[, 2])))
  lsf <- st_sf(geometry = st_sfc(l), crs = 2056)
  lnet <- as_sfnetwork(lsf, directed = FALSE) %E>%
    mutate(weight = edge_length(), highway = highway)
  lnet <- tidygraph::convert(
    lnet,
    to_undirected,
    .clean = TRUE
)
  network <- st_network_join(network, lnet)
}

# Add edges to network
network <- addedge(388, 31782, "footway")
network <- addedge(11152, 11150, "footway")
network <- addedge(13450, 15447, "residential")
network <- addedge(13041, 13131, "residential")
network <- addedge(12490, 12484, "footway")
network <- addedge(13351, 13355, "path")
network <- addedge(13343, 13584, "path")
network <- addedge(25867, 26022, "footway")
network <- addedge(25874, 25871, "footway")
network <- addedge(22794, 22793, "footway")
network <- addedge(33440, 33445, "footway")
network <- addedge(26996, 17023, "residential")
network <- addedge(20545, 20548, "footway")
network <- addedge(18577, 18573, "track")
network <- addedge(18547, 17590, "track")
network <- addedge(16791, 16787, "path")
network <- addedge(16978, 16792, "path")
network <- addedge(17702, 17704, "path")
network <- addedge(2223, 2649, "footway")

network <- network %E>% mutate_if(is.character, ~replace_na(., "unknown"))
network <- network %E>% mutate_if(is.numeric, ~replace_na(., -99))

network <- tidygraph::convert(
    network,
    to_undirected,
    .clean = TRUE
)
```

```{r}
# Split segments with a length >400m into 3 parts for later slope calculation
edges <- network %E>% as_tibble() %>% st_as_sf()

long_edges <- edges %>% filter(drop_units(weight) > 400)
short_edges <- edges %>% filter(drop_units(weight) <= 400)
shortened_edges <- line_segment(long_edges, n_segments = 3)
edges_new <- rbind(short_edges, shortened_edges)
network <- as_sfnetwork(edges_new, directed = FALSE)

network <- network %E>% 
  mutate(
    weight = edge_length())

network <- network %N>%
  mutate(
    degree = centrality_degree(weights = NULL)
  )

network %E>% as_tibble(spatial = FALSE) %>% dplyr::select(weight) %>% summary()
```


## Test accessibility with Badis
```{r}
# Find nearest network node to cell centers. Function from https://github.com/r-spatial/sf/issues/966
find_nearest_node <- function(points, nodes){
  points %>%
  mutate(
    index_of_nearest_feature = st_nearest_feature(., nodes),
    nearest_feature = st_geometry(nodes[index_of_nearest_feature,]),
    nearest_point = purrr::pmap(
      list(geometry, nearest_feature),
      ~ st_nearest_points(.x, .y) %>% st_cast("POINT") %>% magrittr::extract2(2)
    ) 
  ) %>%
  pull(nearest_point) %>%
  st_sfc(crs = 2056)
}

nodes_tibble <- network %N>%
  as_tibble()
nodes_tibble <- st_as_sf(nodes_tibble, sf_column_name = "geometry")

nearest_nodes_pop <- find_nearest_node(statpop, nodes_tibble)
nearest_nodes_badis <- find_nearest_node(badis_zh, nodes_tibble)

# Distances of population points and Badis ZH to nearest network node
summary(st_distance(badis_zh, nearest_nodes_badis, by_element = TRUE))
st_distance(badis_zh, nearest_nodes_badis, by_element = TRUE)

summary(st_distance(statpop, nearest_nodes_pop, by_element = TRUE))
which.max(st_distance(statpop, nearest_nodes_pop, by_element = TRUE))

# Alternatively, blend points onto network: st_network_blend()
```

```{r}
# tmap_mode("view") # set to interactive mode
# tm_tiles("CartoDB.Positron") +
# tm_shape(st_as_sf(network %>% activate("edges")), "edges") +
#   tm_lines() +
# tm_shape(st_as_sf(network %>% activate("nodes"), "nodes")) +
#   tm_dots() +
# tm_shape(badis_zh) + tm_dots(col = "red")
```

```{r}
cost_mat <- st_network_cost(network, from = nearest_nodes_pop, to = nearest_nodes_badis, weights = "weight")
cost_mat <- as.data.frame(cost_mat)
cost_mat <- drop_units(cost_mat)

# Calculate mean accessibility for each population point using an exponential distance weighting function
calcAccessibility <- function(cost_df, dmax, beta){
  
  accessibility <- rep(NA, nrow(cost_df))
  for(i in 1:nrow(cost_df)){
    dists_within_dmax <- cost_df[i, cost_df[i, ] <= dmax] # select those distances below dmax
    dists_within_dmax <- as.numeric(dists_within_dmax)
    imps_within_dmax <- exp(-beta * dists_within_dmax) # exponential distance weighting function
    avg_within_dmax <- mean(imps_within_dmax) # take mean value
    
    accessibility[i] <- avg_within_dmax
    
    accessibility[i] <- ifelse(is.na(accessibility[i]), 0, accessibility[i])
  }
  return(accessibility)
}

statpop["accessibility"] <- calcAccessibility(cost_mat, 5000, 0.00018)
summary(statpop$accessibility)
```

```{r}
statpop <- statpop %>%
  mutate(quantile=ntile(accessibility, 4))

# ggplot() + 
#   geom_sf(data = statpop, aes(col=as.factor(quantile)), size = 0.5) +
#   scale_color_manual(values=c("#132B43", "#28547A", "#3E81B7", "#56B1F7")) +
#   geom_sf(data = badis_zh, aes(fill = "Lidos"), col = "red", shape = 17, size = 2.5) +
#   coord_sf(datum = 2056) +
#   theme_void() +
#   labs(title="Accessibility to Lidos", colour="Accessibility quartile", fill="") +
#   guides(colour = guide_legend(override.aes = list(size=4)))
```

```{r}
# ggplot() + 
#   geom_sf(data = statpop, aes(col=accessibility), size = 0.5) +
#   scale_color_viridis_b() +
#   geom_sf(data = badis_zh, aes(fill = "Badis"), col = "red", shape = 17, size = 2.5) +
#   coord_sf(datum = 2056) +
#   theme_void() +
#   labs(title="Accessibility to Lidos", colour="Accessibility", fill="")
```

```{r}
hist(statpop$accessibility, main = "Distribution of accessibility measure")
```

```{r}
# Look at a random shortest path and compare it to output of Zurich's routing engine
path <- st_network_paths(network, from = statpop[runif(1, min = 1, max = nrow(statpop)), ], to = badis_zh[runif(1, min = 1, max = nrow(badis_zh)), ], weights = "weight")
path

nodes_path <- path %>%
  pull(node_paths) %>%
  unlist() %>% as.numeric()
start_node <- network %N>% st_as_sf() %>% slice(nodes_path[1])
end_node <- network %N>% st_as_sf() %>% slice(nodes_path[length(nodes_path)])
print(c(start_node$geometry, end_node$geometry))

path_net <- network %N>%
  slice(path %>% pull(node_paths) %>% unlist())
print(sum(st_length(st_as_sf(path_net %>% activate("edges")))))

tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(network %>% activate("edges")), "edges") +
  tm_lines(alpha = 0.4) +
tm_shape(st_as_sf(network %>% activate("nodes"), "nodes")) +
  tm_dots(size = 0.02, alpha = 0.4) +
tm_shape(st_as_sf(path_net %>% activate("edges"))) + tm_lines(col = "red") +
tm_shape(st_as_sf(path_net %>% activate("nodes"))) + tm_dots(col = "red", size = 0.01)
```

A few generated shortest paths have been compared to the shortest route of the same points in the city of Zurich network engine. The aim of the thesis is not to generate exact shortes paths (I'm working with an undirected network opposed to the city of Zurich), but the two shortest paths are very similar. More importantly, the lengths of the two paths are almost the same for all generated trips. 

## Save preprocessed network
```{r}
# Save preprocessed network
saveRDS(network, file = here::here("data", "zurich/network_zh.rds"))
edges <- network %E>% as_tibble() %>% st_as_sf()
nodes <- network %N>% as_tibble() %>% st_as_sf()

st_write(edges, dsn = here::here("data", "zurich/edges_zh.shp"), delete_layer = TRUE)
st_write(nodes, dsn = here::here("data", "zurich/nodes_zh.shp"), delete_layer = TRUE)
```

```{r}
# Plot final network
tmap_mode("view") # set to interactive mode
tm_tiles("CartoDB.Positron") +
tm_shape(st_as_sf(network %>% activate("edges")), "edges") +
  tm_lines() +
tm_shape(st_as_sf(network %>% activate("nodes"), "nodes")) +
  tm_dots()
```


